Copulas and their potential for ecology

### 2019-10-06 Paper.Rmd
part way thrugh adding TL stuff to Q3 sections

@@

<!--***DAN: if you do decide to move the TL stuff to main text, this is where the 
methods go. Other than this, I think I am done revising methods for Q3.-->

<!--***DAN: Just ike in the Causes secton you added a comment saying these models
were for initial exploration only and so we did not explore parameter space
nor look at model variants, likewise we also need some commentary in this section 
indicating that the modelling is simple and not comprehensively explored, because
the purpose is just to get an idea of what kinds of consequences of copula 
structure *may* pertain; further work is needed to understand to what extent these
consequences are realized in real systems.-->

Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of the dependences between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is a commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016; @Reuman2017] 
empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and $\log(a)$ is the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We consider that 
Taylor's law is validated for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017].

>>>>

We next addressed Q3 by exploring, using both data and models, three hypothesized
ecological consequences of non-normal copula structure and asymmetric tail
associations. As for Q2, models used are intended as initial explorations, only,
of whether copula structure may possibly have a meaningful influence on important ecological
phenomena. Therefore only simple models are used, and comprehensive explorations of the  
sensitivity of results to model strucuture and parameters are left for future work.

@@

Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of the dependences between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is a commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016; @Reuman2017] 
empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and $\log(a)$ is the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We consider that 
Taylor's law holds true for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017].

<!--***DAN: continung to add TL stuff here, and then next in the following results section-->
